---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r load data}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb")

unzip("activity.zip")
data <- read.csv("activity.csv", header = TRUE)
head(data)
```

## What is mean total number of steps taken per day?

```{r}
library(magrittr)
library(dplyr)

data.by.date <- data %>%
  select(date, steps) %>%
  group_by(date) %>%
  summarize(total.steps = sum(steps)) %>%
  na.omit()

hist(data.by.date$total.steps, xlab = 'Total daily Steps', main = 'Total Steps By Day', breaks = 20)

mean(data.by.date$total.steps)
median(data.by.date$total.steps)
```

## What is the average daily activity pattern?

```{r}
library(ggplot2)
data.by.interval <- data %>%
  select(interval, steps) %>%
  na.omit() %>%
  group_by(interval) %>%
  summarize(total.steps = mean(steps))
ggplot(data.by.interval, aes(x = interval, y = total.steps)) + geom_line()
```
Interval with max number of steps
```{r}
data.by.interval[which(data.by.interval$total.steps == max(data.by.interval$total.steps)), ]
```

## Imputing missing values

```{r}
missing.values <- sum(is.na(data))
missing.values
```
NA values replaced with mean for 5-minute interval
```{r}
library(magrittr)
library(dplyr)

replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
mean.data <- data %>%
  group_by(interval) %>%
  mutate(steps= replacewithmean(steps))
```
Making Histogram and calculating new mean and median
```{r}
alldata.by.day <- aggregate(mean.data$steps, by = list(mean.data$date), sum)

names(alldata.by.day)[1] ="date"
names(alldata.by.day)[2] ="totalsteps"

summary(alldata.by.day)

hist(alldata.by.day$total.steps, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)

mean(alldata.by.day$total.steps)
median(alldata.by.day$total.steps)
```
Mean stayed the same, while median changed. Imputing missing data increased median estimate. This is because NA values, usually set to 0, are replaced with numbers greater than 0.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
mean.data$date <- as.Date(mean.data$date)
mean.data$weekday <- weekdays(mean.data$date)
mean.data$weekend <- ifelse(mean.data$weekday == "Saturday" | mean.data$weekday == "Sunday", "Weekend", "Weekday" )

library(ggplot2)
mean.data.weekend.weekday <- aggregate(mean.data$steps , by = list(mean.data$weekend, mean.data$interval), na.omit(mean))
names(mean.data.weekend.weekday) <- c("weekend", "interval", "steps")

ggplot(mean.data.weekend.weekday, aes(x = interval, y = steps, color = weekend)) + geom_line() + facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean Steps") + ggtitle("Comparison of Average Number of Steps in Each Interval")
```